Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-01 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
working on parspec section

We start by reviewing the ParSpec
framework introduced in Ref.~\cite{Maselli:2019mjd}, but following
some of the notation used in Ref.~\cite{Carullo:2021dui}.

A general procedure to describe deviations from the Kerr QNM frequencies $\omega_{k}^{\rm Kerr}$ and
damping times $\tau_{k}^{\rm Kerr}$ is to write,
%
\begin{subequations}
\label{eq:general_deviation}
\begin{align}
\omega_{k} &= \omega^{\rm Kerr}_{k} \, (1 + \delta \omega_{k}), \\
\tau_{k}   &= \tau^{\rm Kerr}_{k}   \, (1 + \delta \tau_{k}),
\end{align}
\end{subequations}
%
where $\delta\omega_{k}$ and $\delta\tau_{k}$ are the deformation parameters.
%
Here we use the subscript $k$ to represent the QNM labels $\{l,\, m,\, n\}$,
where $l$ and $m$ are the multipole indices, and $n$ the overtone number.
%
This approach was adopted, e.g., in Refs.~\cite{Gossan:2011ha,Meidam:2014jpa,Carullo:2018sfu}.

A caveat of parametrizations of the form~\eqref{eq:general_deviation}, is that
$\delta\omega_{k}$ and $\delta\tau_{k}$ depend, in general, on the BH's source mass and spin.
%
Ideally, one would like to explicitly reinstate the dependence the source-frame
BH mass and spin for it would
%
(i) introduce deformation parameters which can be determined, once and for
all, from a specific gravity theory, and
%
(ii) make it easy to combine constraints from multiple source.

In the ParSpec framework this is accomplished by performing a bivariate
expansion of Eq.~\eqref{eq:general_deviation} in the detector-frame mass and
spin, namely
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{k} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{k} \, \left( 1 + \gamma \delta \omega^{(n)}_{k} \right), \\
\tau_{k}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{k}   \, \left( 1 + \gamma \delta \tau^{(n)}_{k} \right),
\end{align}
\end{subequations}
%
where $M$ and $\chi$ are the detector-frame mass and dimensionless spin of the
black hole; $\omega_{k}^{n}$, $\tau_{k}^{n}$ are dimensionless coefficients
of the spin expansion for Kerr QNMs; $\delta\omega_{k}^{(n)}$, $\delta\tau_{k}^{(n)}$ are
source-independent dimensionless coefficients that characterize the corrections to the Kerr QNM at each spin-order;
and finally $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on the beyond-GR modification is encapsulated in the dimensionless
parameter $\gamma$ given by,
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on the parameter $\ell$, with dimensions of length (and sets the
lengthscale at which beyond-GR modifications become important), and
the exponent $p$ which is related to how the beyond-GR modifications are
included to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made $\ell$ dimensionless by using the other
lenghtscale associated to the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$, which we can also write in terms of the detector-frame mass $M$ through
the redshift $z$~\cite{Krolak:1987ofj}.

>>>>

We start by reviewing the ParSpec
framework introduced in Ref.~\cite{Maselli:2019mjd}, but following
some of the notation used in Ref.~\cite{Carullo:2021dui}.
%
A general procedure to describe deviations to the QNM frequencies $\omega_{k}^{\rm Kerr}$ and
damping times $\tau_{k}^{\rm Kerr}$ of the GR Kerr BH is to write,
%
\begin{subequations}
\label{eq:general_deviation}
\begin{align}
\omega_{k} &= \omega^{\rm Kerr}_{k} \, (1 + \delta \omega_{k}), \\
\tau_{k}   &= \tau^{\rm Kerr}_{k}   \, (1 + \delta \tau_{k}),
\end{align}
\end{subequations}
%
where $\delta\omega_{k}$ and $\delta\tau_{k}$ are the deformation parameters.
%
Here we use the subscript $k$ to represent the QNM labels $\{l,\, m,\, n\}$,
where $l$ and $m$ are the multipole indices, and $n$ the overtone number.
%
This type of parametrization was adopted, for instance, in Refs.~\cite{Gossan:2011ha,Meidam:2014jpa,Carullo:2018sfu}.

A caveat of parametrizations of the form~\eqref{eq:general_deviation}, is that
$\delta\omega_{k}$ and $\delta\tau_{k}$ depend, in general, on the BH's source mass and spin.
%
Ideally, one would like to explicitly reinstate the dependence the source-frame
BH mass and spin for it would
%
(i) introduce deformation parameters which can be determined, once and for
all, from a specific gravity theory (GR included), and
%
(ii) make it simpler to combine constraints coming from multiple source.

In the ParSpec expansion this is accomplished by performing a bivariate
expansion of Eq.~\eqref{eq:general_deviation} in the detector-frame BH's mass
$M$ and spin $\chi$, namely
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{k} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{k} \, \left( 1 + \gamma \delta \omega^{(n)}_{k} \right), \\
\tau_{k}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{k}   \, \left( 1 + \gamma \delta \tau^{(n)}_{k} \right),
\end{align}
\end{subequations}
%
where $\omega_{k}^{n}$ and $\tau_{k}^{n}$ are dimensionless coefficients of the
expansion in spin for Kerr QNMs;
%
$\delta\omega_{k}^{(n)}$ and $\delta\tau_{k}^{(n)}$ are source-independent
dimensionless coefficients that characterize the corrections to the Kerr QNM at
each spin-order;
%
and finally $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on the beyond-GR modification is encapsulated in the dimensionless
parameter $\gamma$ given by,
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on the parameter $\ell$, with dimensions of length (and sets the
lengthscale at which beyond-GR modifications become important), and
the exponent $p$ which is related to how the beyond-GR modifications are
included to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made $\ell$ dimensionless by using the other
lenghtscale associated to the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$, which we can also write in terms of the detector-frame mass $M$ through
the redshift $z$~\cite{Krolak:1987ofj}.
